\documentstyle{amsppt}
\tolerance 3000 \pagewidth{5.5in} \vsize7.0in
\magnification=\magstep1 \NoRunningHeads \NoPageNumbers \topmatter
\title Report to the Executive Committee
\endtitle
\author Alex Iosevich
\endauthor
\endtopmatter
\document

My rankings are as follows \roster \item Kolountzakis
\item Agashe \item Popescu \item Papanikolas \item Saidak \item Milas
\endroster

\vskip.125in

This letter is structured as follows. First, I will make a few
remarks about Mihalis Kolountzakis. I will then comment very
briefly on the rest of the rankings.

\subhead Kolountzakis \endsubhead I would like to echo Pete
Cassaza's sentiments from two years ago. He pointed out during the
Michael Lacey deliberations that we can hire a dozen strong people
without significantly affecting the status of the department. To
move forward, we need to hire high profile individuals respected
by the highest echelons of the mathematical community.

\vskip.125in

Many excellent letters are already in for Mihalis Kolountzakis and
some letters are still on the way. Sergei Konyagin, Paul Cohen,
Noga Alon and Izabella Laba make it crystal clear that Mihalis
Kolountzakis is not just good, but is in fact a first rate
mathematician who proved first rate results in analytic number
theory, computer science, additive number theory, analysis, and
other areas. All of these individuals have excellent credentials
in various aspects of number theory. In addition, Dan Mauldin and
Alexander Olevski comment glowingly on Mihalis' outstanding work
in tiling, an area where number theoretic techniques play a very
significant and frequently a leading role.

\vskip.125in

In addition to his deep knowledge of number theory and analysis,
Mihalis Kolountzakis is a respected expert in the use of
probabilistic methods. He is cited in a celebrated book by Alon
and Spencer, entitled "Probabilistic method" as one of the
pioneers of de-randomization, a method of obtaining deterministic
results from probabilistic methods using conditional probabilities
and computer algorithms. Mihalis has successfully applied these
techniques to solve interesting problems in additive number
theory, analytic number theory, and other areas.

\vskip.125in

\subhead Other candidates \endsubhead After re-examining the
files, I have come to the conclusion that as promising as Saidak
appears to be, he does not compare at this stage of his career
with the other candidates on the short list, except, possibly,
Milas. I cannot, in good conscience, rate him above Agashe,
Popescu and Papanikolas, even though he is closer to my number
theoretic research interests than all the other candidates except
Kolountzakis.

Agashe is also somewhat young, but his promise and qualifications
are striking enough that I rate him second for this position,
above everyone except Mihalis Kolountzakis who I believe is a
super-star and an ideal candidates for this position.










\enddocument
